---
title: "Class description of TPenMonteith"
author: "Agronomy and Crop Science Group, Kiel University"
date: "`r Sys.Date()`"
bibliography: bibliography.bibtex.bib
csl: european-journal-of-agronomy.csl
chunk_output_type: console
html-math-method: katex
number-sections: true
toc: true
toc-expand: 4
toc-depth: 4
toc-location: left
crossref:
  eq-prefix: Eq.
format:
 html:
  mermaid:
    theme: forest
    mainfont: Helvetica Neue, Helvetica, Arial, sans
    monofont: Cascadia Mono, Menlo, Consolas, Courier New, Courier
    mathfont: "TeX Gyre Termes Math"
---

```{r}
#| label: SetupLibraries
#| include: FALSE

knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
rm(list = ls(all.names = TRUE))

library(rmarkdown)
library(bookdown)
library(soilwater)
library(pracma)
library(dplyr)
library(ggplot2)
library(extrafont)
library(gifski)
library(gganimate)
library(transformr)
library(reticulate)
library(tinytex)
library(magick)
library(bibtex)
library(ggsci)
library(knitcitations)
library(kableExtra)

options("citation_format" = "pandoc")

```



# Introduction

The class TPenMonteith calculates the potential evapotranspiration according to the Penman-Monteith equation. The Penman-Monteith equation is a combination of the Penman equation and the Monteith equation. The Penman equation calculates the evaporation from a free water surface, while the Monteith equation calculates the transpiration from a plant. The Penman-Monteith equation calculates the potential evapotranspiration, which is the sum of the evaporation and the transpiration. The potential evapotranspiration is the evapotranspiration that would occur if there were no water limitation.


## Descendence

The class TPenMonteith is derived from TPlantRelatedSubMod, which is derived from TSubmodel which is derived from TObject or TGraphicControl.

TPenMonteith
           |____TPlantRelatedSubMod
                                  |____TSubmodel


## State variables

```{r}
#| message: false
#| warning: false
#| include: false
fn <- "Q:/OSR/INI/ModelLowN/Docu_Docu_ModelLown_talent_N150_Neff.csv"
df <- read.delim(fn, header = TRUE, sep = ";")
df.state <- df %>% filter(Submodel == "PenMonteith1", EntityType == "State") %>% dplyr::select(EntityName, Units, Value, Comment)
names(df.state) <- c("State variable", "Units", "InitialValue", "Description")
```
The class TPenMonteith has `r trunc(nrow(df.state))`  following state variable(s).


```{r}
#| echo: false
#| message: false
#| warning: false
kable(df.state)
```

## Parameters

```{r}
#| message: false
#| warning: false
#| include: false

df.par <- df %>% filter(Submodel == "PenMonteith1", EntityType == "Parameter") %>% dplyr::select(EntityName, Units, Value, Comment)
names(df.par) <- c("Parameter", "Units", "Value", "Description")

```

The class TPenMonteith has `r trunc(nrow(df.par))`  following parameter(s).

```{r}
#| echo: false
#| message: false
#| warning: false
kable(df.par,  escape = FALSE)
```
## Variables

```{r}
#| message: false
#| warning: false
#| include: false

df.var <- df %>% filter(Submodel == "PenMonteith1", EntityType == "Variable") %>% dplyr::select(EntityName, Units, Comment)
names(df.var) <- c("Variable", "Units", "Description")

```

The class TPenMonteith has `r trunc(nrow(df.var))`  following variable(s).

```{r}
kable(df.var,  escape = FALSE)
```



## External variables

```{r}
#| message: false
#| warning: false
#| include: false

df.ext <- df %>% filter(Submodel == "PenMonteith1", EntityType == "ExternalValue") %>% dplyr::select(EntityName, Units, Comment, Option)
names(df.ext) <- c("External variable", "Units", "Description", "Source")

```

The class TPenMonteith has `r trunc(nrow(df.ext))`  following external variable(s).

```{r}
#| echo: false
#| message: false
#| warning: false
kable(df.ext,  escape = FALSE)
```


## Options

```{r}
#| message: false
#| warning: false
#| include: false

df.opt <- df %>% filter(Submodel == "PenMonteith1",EntityType == "Option" ) %>% dplyr::select(EntityName, Units, Comment)
names(df.opt) <- c("Option", "Units", "Description")

```

The class TPenMonteith has `r trunc(nrow(df.opt))`  following option(s).


```{r}
#| echo: false
#| message: false
#| warning: false
kable(df.opt,  escape = FALSE)
```
