---
title: "TSoilWaterMod"
author: "Agronomy and Crop Science Group, Kiel University"
date: "`r Sys.Date()`"
chunk_output_type: console
html-math-method: katex
number-sections: true
toc: true
toc-expand: 4
toc-depth: 4
toc-location: left
crossref:
  eq-prefix: Eq.
format: 
  html:
    embed-resources: true 
---

```{r}
#| label: SetupLibraries
#| include: FALSE

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center", fig.width = 12, fig.height = 6, fig.pos = "h", fig.cap = TRUE)
rm(list = ls(all.names = TRUE))

library(rmarkdown)
library(bookdown)
library(soilwater)
library(pracma)
library(dplyr)
library(ggplot2)
library(extrafont)
library(gifski)
library(gganimate)
library(transformr)
library(reticulate)
library(tinytex)
library(magick)
library(bibtex)
library(ggsci)
library(knitcitations)
library(kableExtra)
library(Weatherfunctions)
library(xml2)
library(ggpmisc)
library(tidyr)
source("Q:/HUME/HUME/RLib/DokuUtilFunctions.R")

options("citation_format" = "pandoc")

```

```{r}
#| include: false
fn_source <- "Q:/HUME/HUME/Components/Soil/URootedSoil.pas"
fn_xml_docu <- "Q:/HUME/HUME/XML_Delphi_Docu/URootedSoil.xml"
```

# Introduction

```{r}
#| echo: false
#| results: "asis"
#| message: false
#| warning: false

# 1. Load the XML file
doc <- xml2::read_xml(fn_xml_docu)


```

```{r}
#| echo: false
#| results: "asis"
#| message: false
#| warning: false

GetClassDevNotes("TSoilWaterModelR",fn_xml_docu)
```

## Ancestor classes

The class TSoilWaterModelR has the following ancestor classes shown in Table @tbl-AncestorClasses.



```{r}
#| label: tbl-AncestorClasses
#| echo: false
#| tbl-cap: "Ancestor classes of TSoilWaterModelR."

df <- GetAncestorClasses(fn_xml_docu,"TSoilWaterModelR")

kable(df)



```


# Scientific Background

## Soil water transport

## Root water uptake

## Soil evaporation

Actual evaporation \[mm d^-1^\] (E~act~) is calculated by a reduction factor (f~evap~), which corrects E~pot~ by the influence of low soil water content in the top soil layer.

$${{E}_{act}}={{E}_{pot}}\cdot {{red}_{evap}}$$ {#eq-Eact}

Within the module two different options for calculation of evaporation are included. The option called *modified Beese* is the default:

$${{red}_{evap}}=max\left(0, min \left(1,\frac{-1\cdot (\log 10(ps{{i}_{arr[1]}})-4.2)}{4.2-\log 10(ps{{i}_{critevap}})}\right)\right)$$ {#eq-fevap}

the original function according to Beese1978 however is available as well:

$$red_{evap} = max\left(0, min \left(1,-0.5767 * log_{10}(psi_{arr1}) + 1.78\right)\right)$$ {#eq-fevapBeese}

Thereby psi~arr~\[1\] is the soil water tension in the top soil compartment and psi~critevap~ is the soil water tension at which evaporation is reduced. The initiation of soil water potential (psi) is simply done by a parameter psi~ini~. The parameter value of psi~critevap~ has a default value of 500 \[hPa\].

```{r}
#| echo: false
#| results: "asis"
#| message: false
#| warning: false
#| fig.cap: "Reduction factor for soil evaporation (f~evap~) as a function of soil water tension in the top soil layer (psi~arr~[1]) for a given critical soil water tension for evaporation reduction (psicritevap)."



psi_arr_1 <- seq(-14000, -1, by = 1) # Soil water tension in the top soil layer

psicritevap <- -500 # Critical soil water tension for evaporation reduction
fevap <- length(psi_arr_1)
fevap <- pmax(0, pmin(1,(-1 * (log10(-psi_arr_1) - 4.2)) / (4.2 - log10(-psicritevap))))
fevapBeese <- pmax(0, pmin(1, -0.5767 * log10(-psi_arr_1) + 1.78))


illu <- data.frame(psi_arr_1, fevap, fevapBeese)
illu <- pivot_longer(illu, cols = c("fevap", "fevapBeese"), names_to = "Method", values_to = "ReductionFactor")


p <- ggplot(data = illu) +
  geom_line( aes(x = abs(psi_arr_1), y = ReductionFactor, group=Method, color=Method), size = 1) +
  scale_x_log10() +
  scale_color_manual(values = c("blue", "red"), labels = c("Modified Beese", "Original Beese")) +
  labs(x = "Soil water tension in the top soil layer (psi_arr[1]) [kPa]",
       y = "Reduction factor for soil evaporation (fevap)") +
  theme_minimal()
p

```
