---
title: "TAbstractPlant"
author: "Agronomy and Crop Science Group, Kiel University"
date: "`r Sys.Date()`"
chunk_output_type: console
html-math-method: katex
number-sections: true
toc: true
toc-expand: 4
toc-depth: 4
toc-location: left
format:
  html:
    embed-resources: true
---

```{r}
#| label: setup-libraries
#| include: false

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(dplyr)
library(kableExtra)
library(knitr)
```

```{r}
#| label: setup-documentation-data
#| include: false

source_file <- "Components/PlantComponents/UAbstractPlant.pas"
documentation_csv <- "csv_Hume_Docu/GrowthCurvePlant1.csv"

if (!file.exists(source_file)) {
  stop("Source file not found: ", source_file)
}

if (!file.exists(documentation_csv)) {
  stop(
    "Model-element documentation CSV not found: ", documentation_csv,
    ". Generate it with TMod.write_documentation and copy it to csv_Hume_Docu."
  )
}

entity_data <- read.delim(
  documentation_csv,
  header = TRUE,
  sep = ";",
  fileEncoding = "UTF-8-BOM",
  check.names = TRUE,
  stringsAsFactors = FALSE
)

# Older documentation CSV files used "SubModel" and "ExtVar". Normalize them
# so that this draft works with both the legacy and current output formats.
if ("SubModel" %in% names(entity_data) &&
    !"Submodel" %in% names(entity_data)) {
  names(entity_data)[names(entity_data) == "SubModel"] <- "Submodel"
}

entity_data <- entity_data %>%
  mutate(EntityType = recode(EntityType, ExtVar = "ExternalValue"))

# The checked-in example predates registration-time provenance. These explicit
# base-class lists also make the inheritance shown in this document independent
# of the concrete representative plant class.
submodel_entities <- c("ContOutput", "FinalOutput")
abstract_plant_entities <- c(
  "C_Residues", "N_Residues", "DMtotal", "DMfineroot",
  "SowingDate", "HarvestDate", "Par_Psi2", "Par_rc0",
  "Par_ExtCoeffPAR", "Par_Weff", "WithRoots"
)

if (!"DeclarationClass" %in% names(entity_data)) {
  entity_data$DeclarationClass <- "TGrowthCurvePlant"
}

entity_data <- entity_data %>%
  mutate(
    IntroductionClass = case_when(
      EntityName %in% submodel_entities ~ "TSubmodel",
      EntityName %in% abstract_plant_entities ~ "TAbstractPlant",
      TRUE ~ DeclarationClass
    )
  )

# Documentation output can contain one copy per INI file. Use one scenario for
# structural entity tables because class provenance does not vary by scenario.
if ("IniFile" %in% names(entity_data)) {
  first_ini_file <- unique(entity_data$IniFile)[1]
  entity_data <- entity_data %>% filter(IniFile == first_ini_file)
}

entity_data <- entity_data %>% distinct()

make_entity_table <- function(entity_type, columns, headings) {
  result <- entity_data %>%
    filter(EntityType %in% entity_type) %>%
    select(IntroductionClass, all_of(columns)) %>%
    distinct()

  names(result) <- c("Introduced in class", headings)
  result
}

state_table <- make_entity_table(
  "State",
  c("EntityName", "Units", "Value", "Comment"),
  c("State variable", "Units", "Initial value", "Description")
)

parameter_table <- make_entity_table(
  "Parameter",
  c("EntityName", "Units", "Value", "Comment"),
  c("Parameter", "Units", "Value", "Description")
)

variable_table <- make_entity_table(
  "Variable",
  c("EntityName", "Units", "Comment"),
  c("Variable", "Units", "Description")
)

constant_table <- make_entity_table(
  "Constant",
  c("EntityName", "Units", "Value", "Comment"),
  c("Constant", "Units", "Value", "Description")
)

external_table <- make_entity_table(
  "ExternalValue",
  c("EntityName", "Units", "Option", "Comment"),
  c("External variable", "Units", "Source", "Description")
)

option_table <- make_entity_table(
  "Option",
  c("EntityName", "Option", "Comment"),
  c("Option", "Default value", "Description")
)

render_entity_table <- function(table_data, caption) {
  if (nrow(table_data) == 0) {
    cat("No entities of this type are present in the documentation data.")
  } else {
    print(
      kable(
        table_data,
        caption = caption,
        escape = FALSE
      ) %>%
        kable_styling(
          bootstrap_options = c("striped", "hover", "condensed"),
          full_width = TRUE
        )
    )
  }
}
```

# Summary

`TAbstractPlant` is the abstract base class for HUME plant components. It provides common crop states and parameters, crop-rotation links to related soil and evapotranspiration submodels, and the interfaces that concrete plant classes must implement for leaf area, crop height, roots, dry-matter growth, and nitrogen uptake.

The same source unit also declares `TPlantRelatedSubMod`, the base class for submodels that maintain a link to the currently active plant in a crop rotation.

Source unit: `Components/PlantComponents/UAbstractPlant.pas`

## Class hierarchy

``` text
TSubmodel
|-- TPlantRelatedSubMod
`-- TAbstractPlant
    `-- concrete plant models, for example TGrowthCurvePlant
```

| Class | Responsibility |
|------------------------------------|------------------------------------|
| `TPlantRelatedSubMod` | Stores the active plant model, remembers the first linked plant, and restores that link during initialization. |
| `TAbstractPlant` | Defines common plant entities, crop-rotation behavior, harvest handling, and the abstract plant/soil coupling interface. |

## Source provenance

The tables below use registration-time class information from the `DeclarationClass` field generated by `TMod.write_documentation`. The entity lists declared directly by `TSubmodel` and `TAbstractPlant` are also stated explicitly in the setup chunk. This ensures that inherited entities are assigned to their introducing base class even when the representative CSV was generated from a concrete descendant.

The checked-in `GrowthCurvePlant1.csv` still uses the older CSV layout. It therefore receives `TGrowthCurvePlant` as the default class, after which the known base-class entities are assigned to `TSubmodel` and `TAbstractPlant`. Regenerating the CSV with the current code supplies the stored RTTI provenance for all other entities.

# Model entities

The tables describe the complete entity set of the representative `TGrowthCurvePlant` instance. The first column identifies the class that introduced each entity.

## State variables

The representative plant instance contains `r nrow(state_table)` state variable(s).

```{r}
#| label: table-state-variables
#| results: asis
render_entity_table(state_table, "Plant state variables")
```

## Parameters

The representative plant instance contains `r nrow(parameter_table)` parameter(s).

```{r}
#| label: table-parameters
#| results: asis
render_entity_table(parameter_table, "Plant parameters")
```

## Variables

The representative plant instance contains `r nrow(variable_table)` calculated variable(s).

```{r}
#| label: table-variables
#| results: asis
render_entity_table(variable_table, "Calculated plant variables")
```

## Constants

The representative plant instance contains `r nrow(constant_table)` registered constant(s).

```{r}
#| label: table-constants
#| results: asis
render_entity_table(constant_table, "Plant constants")
```

## External variables

The representative plant instance contains `r nrow(external_table)` external variable(s).

```{r}
#| label: table-external-variables
#| results: asis
render_entity_table(external_table, "External plant variables")
```

## Options

The representative plant instance contains `r nrow(option_table)` option(s).

```{r}
#| label: table-options
#| results: asis
render_entity_table(option_table, "Plant options")
```

# Common behavior

## Initialization

`TAbstractPlant.CreateAll` registers the dates of sowing and harvest, residue pools, total and fine-root dry matter, coupling parameters for water uptake and evapotranspiration, and the `WithRoots` option. During initialization, `WithRoots` is converted to the internal Boolean flag and the residue pools and harvest state are reset.

## Harvest and crop rotation

`TAbstractPlant.Integrate` performs harvest when the configured harvest date is passed or `DoHarvest` is set. It transfers carbon and nitrogen residues to a connected soil-mineralization model, activates the next crop, and transfers the linked soil-water, soil-nitrogen, soil-layer, and evapotranspiration submodels to that crop. If `SetNewDates` is enabled, sowing and harvest dates are advanced by `RotationLength` years.

## Interface for concrete plant models

Concrete descendants must provide access to:

- leaf area index and crop height;
- nitrogen uptake rate;
- root-length density by soil layer;
- total and effective root length; and
- the dry-matter growth rate.

The common parameters `Par_ExtCoeffPAR`, `Par_rc0`, `Par_Psi2`, and `Par_Weff` allow other HUME components to obtain plant-specific radiation, canopy, water-stress, and rooting-depth values through the abstract plant interface.

# Draft status

This document is a source-code-based first draft. Before publication, regenerate the representative CSV with the current `TMod.write_documentation`, review the descriptions and units emitted by the model, and remove the legacy CSV fallback once the checked-in example data has been updated.
