---
title: "TGrowthCurvePlant"
author: "Agronomy and Crop Science Group, Kiel University"
date: "`r Sys.Date()`"
bibliography: bibliography.bibtex.bib
csl: european-journal-of-agronomy.csl
chunk_output_type: console
html-math-method: katex
number-sections: true
toc: true
toc-expand: 4
toc-depth: 4
toc-location: left
crossref:
  eq-prefix: Eq.
format: 
  html:
    embed-resources: true 
---

```{r}
#| label: SetupLibraries
#| include: FALSE

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center", fig.width = 12, fig.height = 8, fig.pos = "h", fig.cap = TRUE)
rm(list = ls(all.names = TRUE))

library(rmarkdown)
library(bookdown)
library(soilwater)
library(pracma)
library(dplyr)
library(ggplot2)
library(extrafont)
library(gifski)
library(gganimate)
library(transformr)
library(reticulate)
library(tinytex)
library(magick)
library(bibtex)
library(ggsci)
library(knitcitations)
library(kableExtra)
library(Weatherfunctions)
library(xml2)
library(ggpmisc)
library(tidyr)
options("citation_format" = "pandoc")

```

```{r}
#| label: SetupPathsReadXML
#| message: false
#| warning: false
#| include: false
fn_source <- "Q:/HUME/HUME/Components/PlantComponents/UGrowthCurvePlant.pas"
fn_xml_docu <- "Q:/HUME/HUME/XML_Delphi_Docu/UGrowthCurvePlant.xml"
doc <- read_xml(fn_xml_docu)


```

# Summary

```{r}
#| label: SummaryDevnotes
#| echo: false
#| message: false
#| warning: false

classes <- xml_find_all(doc, ".//class")

# Extract <devnotes> content and class names
result <- data.frame(
  class_name = xml_attr(classes, "name"),
  devnotes   = sapply(classes, function(class_node) {
    devnotes_node <- xml_find_first(class_node, "./devnotes")
    if (!is.na(devnotes_node)) trimws(xml_text(devnotes_node)) else NA_character_
  }),
  stringsAsFactors = FALSE
)
```

```{r}
# Find the TGrowthCurvePlant class node
class_node <- xml_find_first(doc, ".//class[@name='TGrowthCurvePlant']")

# Extract all propertyref nodes inside that class
properties <- xml_find_all(class_node, ".//propertyref")

# Extract attributes: name and visibility
property_df <- data.frame(
  name       = xml_attr(properties, "name"),
  visibility = xml_attr(properties, "visibility"),
  stringsAsFactors = FALSE
)
```

```{r}
library(xml2)
library(dplyr)
library(tidyr)
library(stringr)

# Load XML
#doc <- read_xml("UGrowthCurvePlant.xml")

# Find TGrowthCurvePlant class node
class_node <- xml_find_first(doc, ".//class[@name='TGrowthCurvePlant']")

# Extract <propertyref> elements
property_refs <- xml_find_all(class_node, ".//propertyref")

property_df <- data.frame(
  name       = xml_attr(property_refs, "name"),
  visibility = xml_attr(property_refs, "visibility"),
  stringsAsFactors = FALSE
)



# Extract <field> elements under <members>
#field_nodes <- xml_find_all(class_node, ".//members/field")
field_nodes <- xml_find_all(class_node, ".//field")

# Extract name, type, and optional <summary> and <remarks>
field_df <- data.frame(
  name     = xml_attr(field_nodes, "name"),
  type     = xml_attr(field_nodes, "type"),
  summary  = sapply(field_nodes, function(x) {
    summary_node <- xml_find_first(x, ".//summary")
    if (!is.na(summary_node)) xml_text(summary_node) else NA_character_
  }),
  remarks  = sapply(field_nodes, function(x) {
    remarks_node <- xml_find_first(x, ".//remarks")
    if (!is.na(remarks_node)) xml_text(remarks_node) else NA_character_
  }),
  stringsAsFactors = FALSE
)

# Combine property_ref and field info by matching name
full_df <- left_join(property_df, field_df, by = "name")

# Optional: clean up whitespace
full_df <- full_df %>%
  mutate(across(where(is.character), ~str_squish(.)))

# View result
#print(full_df)

```

```{r}
#| label: SummaryDevnotesOutput
#| echo: false
#| message: false
#| warning: false
#| results: asis
#cat(result$devnotes, sep="\n\n")

```

`r result$devnotes`

## Descendence

The class TGrowthCurvePlant is derived from TAbstractPlant which derived from TSubmodel which is derived from TObject or TGraphicControl when compiled for a console app or a Delphi VCL application.

TAbstractPlant \|\_\_\_\_TSubmodel

## State variables

```{r}
#| label: StateVariables
#| message: false
#| warning: false
#| include: false
#fn <- "Q:/SimProject/project/GrowthCurvePlant1.csv"
fn <- "csv_Hume_Docu/GrowthCurvePlant1.csv"
df <- read.delim(fn, header = TRUE, sep = ";")
df.state <- df %>% filter(SubModel == "GrowthCurvePlant1", EntityType == "State") %>% dplyr::select(EntityName, Units, Value, Comment)
names(df.state) <- c("State variable", "Units", "InitialValue", "Description")
```

The class TGrowthCurvePlant has `r trunc(nrow(df.state))` following state variable(s).

```{r}
#| label: StateVariablesOutput
#| echo: false
#| message: false
#| warning: false
kable(df.state)
```

## Parameters

```{r}
#| label: Parameters
#| message: false
#| warning: false
#| include: false

df.par <- df %>% filter(SubModel == "GrowthCurvePlant1", EntityType == "Parameter") %>% dplyr::select(EntityName, Units, Value, Comment)
names(df.par) <- c("Parameter", "Units", "Value", "Description")

```

The class TGrowthCurvePlant has `r trunc(nrow(df.par))` following parameter(s).

```{r}
#| label: ParametersOutput
#| echo: false
#| message: false
#| warning: false
kable(df.par,  escape = FALSE)
```

## Variables

```{r}
#| label: Variables
#| message: false
#| warning: false
#| include: false

df.var <- df %>% filter(SubModel == "GrowthCurvePlant1", EntityType == "Variable") %>% dplyr::select(EntityName, Units, Comment)
names(df.var) <- c("Variable", "Units", "Description")

```

The class TGrowthCurvePlant has `r trunc(nrow(df.var))` following variable(s).

```{r}
#| label: VariablesOutput
#| echo: false
#| message: false
#| warning: false
kable(df.var,  escape = FALSE)
```

## External variables

```{r}
#| label: ExternalVariables
#| message: false
#| warning: false
#| include: false

df.ext <- df %>% filter(SubModel == "GrowthCurvePlant1", EntityType == "ExtVar") %>% dplyr::select(EntityName, Units, Comment, Option)
names(df.ext) <- c("External variable", "Units", "Description", "Source")

```

The class TGrowthCurvePlant has `r trunc(nrow(df.ext))` following external variable(s).

```{r}
#| label: ExternalVariablesOutput
#| echo: false
#| message: false
#| warning: false
kable(df.ext,  escape = FALSE)
```

## Options

```{r}
#| label: Options
#| message: false
#| warning: false
#| include: false

df.opt <- df %>% filter(SubModel == "GrowthCurvePlant1",EntityType == "Option" ) %>% dplyr::select(EntityName, Units, Comment)
names(df.opt) <- c("Option", "Units", "Description")

```

The class TGrowthCurvePlant has `r trunc(nrow(df.opt))` following option(s).

```{r}
#| label: OptionsOutput
#| echo: false
#| message: false
#| warning: false
kable(df.opt,  escape = FALSE)
```

## Scientific Background

The interpolation of measured values is a task which is supported by the class TGrowthCurvePlant. Thereby growth curves as well as linear interpolation can be used.

### Growth curves

The growth curves supported by this class are:

-   Linear Growth
-   Exponential growth
-   Logistic growth
-   Expolinear growth
-   Richards function
-   Logistic with decay
-   LogInt growth
-   LogInt with decay

### Temperature sum and effective Temperature

All growth curves are based on growth rates which are defined as a function of the effectiv temperature $T_{eff}$, which is the daily mean temperatures above a base temperature $T_b$.

The effective temperature is defined as:

$$ T_{eff} = max \left( 0, T_{mean}-T_b  \right) $$ {#eq-effective-temperature}

where $T_{mean}$ is the daily mean temperature and $T_b$ is the base temperature.

The temperature sum $T_{sum}$ is defined as the sum of the effective temperatures over time:

$$ T_{sum} = \sum_0^t T_{eff}(t) dt $$ {eq -temperature-sum}

where $T_{sum}$ is the temperature sum at time $t$ and $T_{eff}(t)$ is the effective temperature at time $t$.

#### Linear growth

The linear growth curve is defined by the following equation:

$$ \frac{dW}{dt} = gr \cdot T_{eff} $$ {#eq-linear-growth}

where $dW/dt$ is the growth rate and $gr$ is the growth rate parameter.

Integration with respect to temperature sum and with W_0 as the initial value gives:

$$ W(t) = W_0 + gr \cdot Tsum $$ {#eq-linear-growth-integration}

where $W(t)$ is the weight at temperature sum $T_{sum}$ and $W_0$ is the initial weight.

#### Exponential growth

The exponential growth curve is defined by the following equation:

$$ \frac{dW}{dt} = rgr \cdot T_{eff} \cdot W $$ {#eq-exponential-growth}

where $dW/dt$ is the growth rate, $rgr$ is a growth rate parameter and $W$ is the weight.

Integration with respect to temperature sum and with $W_0$ as the initial value gives:

$$ W(t) = W_0 \cdot e^{rgr \cdot T_{sum}} $$ {#eq-exponential-growth-integration}

where $W(t)$ is the weight at temperature sum $T_{sum}$ and $W_0$ is the initial weight.

#### Logistic growth

The logistic growth curve is defined by the following equation:

$$ \frac{dW}{dt} = rgr \cdot T_{eff} \cdot W \cdot (1 - \frac{W}{W_{max}}) $$ {#eq-logistic-growth}

where $dW/dt$ is the growth rate, $rgr$ is a growth rate parameter, $W$ is the weight and $W_{max}$ is the maximum weight.

Integration with respect to temperature sum and with $W_0$ as the initial value gives:

$$ W(t) = \frac{W_{max} \cdot W_0}{W_{max} + (W_{0} - W_{max}) \cdot e^{-rgr \cdot T_{sum}}} $$ {#eq-logistic-growth-integration}

where $W(t)$ is the weight at temperature sum $T_{sum}$ and $W_0$ is the initial weight.

#### Expolinear growth

The expolinear growth curve is defined by the following equation:

$$ \frac{dW}{dt} = rgr \cdot T_{eff} \cdot W \cdot (1 - e^{-k \cdot T_{sum}}) $$ {#eq-expolinear-growth}

where $dW/dt$ is the growth rate, $rgr$ is a growth rate parameter, $W$ is the weight, $k$ is a decay parameter and $T_{sum}$ is the temperature sum.

Integration with respect to temperature sum and with $W_0$ as the initial value gives:

$$ W(t) = W_0 \cdot e^{rgr \cdot T_{sum}} \cdot (1 - e^{-k \cdot T_{sum}}) $$ {#eq-expolinear-growth-integration}

where $W(t)$ is the weight at temperature sum $T_{sum}$ and $W_0$ is the initial weight.

#### Richards function

The Richards function is defined by the following equation:

$$ \frac{dW}{dt} = rgr \cdot T_{eff} \cdot W \cdot (1 - (W/W_{max})^{\frac{1}{m}}) $$ {#eq-richards-growth}

where $dW/dt$ is the growth rate, $rgr$ is a growth rate parameter, $W$ is the weight, $W_{max}$ is the maximum weight and $m$ is a shape parameter.

Integration with respect to temperature sum and with $W_0$ as the initial value gives:

$$ W(t) = W_{max} \cdot (1 + (W_{0}/W_{max})^{m} \cdot e^{-rgr \cdot T_{sum}})^{-\frac{1}{m}} $$ {#eq-richards-growth-integration}

where $W(t)$ is the weight at temperature sum $T_{sum}$ and $W_0$ is the initial weight.

#### Logistic with decay

The logistic with decay growth curve is defined by the following equation:

$$ \frac{dW}{dt} = rgr \cdot T_{eff} \cdot W \cdot (1 - \frac{W}{W_{max}}) - k \cdot W $$ {#eq-logistic-decay-growth}

where $dW/dt$ is the growth rate, $rgr$ is a growth rate parameter, $W$ is the weight, $W_{max}$ is the maximum weight and $k$ is a decay parameter.
