---
title: "TSimpleRootModDM"
author: "Agronomy and Crop Science Group, Kiel University"
date: "`r Sys.Date()`"
chunk_output_type: console
html-math-method: katex
number-sections: true
toc: true
toc-expand: 4
toc-depth: 4
toc-location: left
crossref:
  eq-prefix: Eq.
format: 
  html:
    embed-resources: true 
---

```{r}
#| label: SetupLibraries
#| include: FALSE

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center", fig.width = 12, fig.height = 6, fig.pos = "h", fig.cap = TRUE)
rm(list = ls(all.names = TRUE))

library(rmarkdown)
library(bookdown)
library(soilwater)
library(pracma)
library(dplyr)
library(ggplot2)
library(extrafont)
library(gifski)
library(gganimate)
library(transformr)
library(reticulate)
library(tinytex)
library(magick)
library(bibtex)
library(ggsci)
library(knitcitations)
library(kableExtra)
library(Weatherfunctions)
library(xml2)
library(here)
library(ggpmisc)
library(tidyr)
source("RLib/DokuUtilFunctions.R")

options("citation_format" = "pandoc")

```

```{r}
#| include: false
fn_source <- "Components/Rootgrowth/USimpleRootModDM.pas"
fn_xml_docu <- "XML_Delphi_Docu/USimpleRootModDM.xml"
```

# Summary

```{r}
#| echo: false
#| results: "asis"
#| message: false
#| warning: false

# 1. Load the XML file
doc <- xml2::read_xml(fn_xml_docu)


```

```{r}
#| echo: false
#| results: "asis"
#| message: false
#| warning: false

GetClassDevNotes("TSimpleRootModDM",fn_xml_docu)
```

## Ancestor classes

The class TSimpleRootModDM has the following ancestor classes shown in @tbl-AncestorClasses.

```{r}
#| label: tbl-AncestorClasses
#| echo: false
#| tbl-cap: "Ancestor classes of TSimpleRootModDM"

df <- GetAncestorClasses(fn_xml_docu,"TSimpleRootModDM")

kable(df)

```

# Scientific Background

## Root length and root length density

In order to estimate the water transport rate towards roots and their possible limitations it is necessary to have a realistic estimate of the total root length, its spatial an temporal variation and furthermore the distribution of water uptake over the parts of the root system. The latter point is often governed by the variation of water tension within the rooted soil layers, as during a drying phase the more densely rooted upper soil layers dry out and the water uptake than focuses on deeper soil layers.

The variation total root length of a crop over time varies between crops and for a specific crop total root length is somewhat related to the total biomass production and its influencing factors. Cereal crops may have maximum total root root length values of 10 $km^2/m^2$ or in other units 100 $[cm\ cm^{-2}]$. Especially grain legumes have much lower total root length values, down to approximately 10% of the values of cereal crops, which may lead to a value of 10 $[cm\ cm^{-2}]$ [@Kage1996d]. For a rooting depth of 100 cm (cereal crop) this leads to an average root length density of 1 $[cm \cdot cm^{-3}]$ and for a lower rooting depth of a grain legume of about 50 \[cm\] we end at values of 0.2 $[cm\ cm^{-2}]$. Root length density in the subsoil, which is most important during drought stress phases, however, is much lower.

## Dynamic of total root length

The growth of the root system is closely linked to the growth of the entire crop and to some extend and under some conditions (drought stress, nutrient deficiency) *vice* *versa*. If the fraction of dry matter increase attributed to fine root growth, ffR, is known, the increase of total fine root length dRL/dt (cm.m-2.d-1) may simply be calculated from the total dry weight increase dWt/dt $(g \cdot m^{-2} \cdot d^{-1)}$ and the average specific root length SRL (sp_WL) $(cm\cdot g^{-1} DM)$ [@Kage2000].

$$\frac{dRL}{dt}=\frac{d{{W}_{t}}}{dt}\cdot {{f}_{fR}}\cdot SRL $$ {#eq-dRLdt}

The value of $f_{fR}$ may be assumed to be constant or to decrease during the plant’s development from the vegetative to the generative phase with a minimum value of zero. As a simple approximation a linear decrease of $f_{fR}$ with temperature sum since emergence may be assumed:

$$f_{fR}=\max (0,\,\,{{f}_{fR0}}-{{f}_{fRdec}}\cdot TS)$$ {#eq-ffR}

where f~fR0~ (-) is the initial fraction of dry matter allocated to the root fraction and f~fRdec~ $(°C^{-1} d^{-1})$ is the decrease of this fraction per unit of accumulated temperature. The value of SRL may regarded as a parameter or be calculated from the average diameter of the roots and the average dry matter content of the roots. A constant value of 7000 $cm\cdot g^{-1}\; DM$ is used here based on data of @Barraclough1984, nevertheless, also much lower values have been reported from soils with a quite high clay content [@Siddique1990; @Savin1994].

## Shoot dry matter production and total dry matter increase

The total dry matter production rate may be estimated from the shoot dry matter increase if one considers only a fine root pool as below-ground biomass:

$$\frac{dW_t}{dt}=\frac{1}{\left( 1-{{f}_{fR}} \right)}\frac{d{{W}_{sh}}}{dt}$$ {#eq-dWtdt}

Shoot dry matter increase may be obtained from a crop growth model or may be derived from an appropriate function fitted to experimental data. Here, as an example an approach, using the Richards-function [@Thornley1990] is shown:

$$\frac{dW_{sh}}{dt}=r_{Ws}\cdot T_{eff} \cdot W_{sh} \cdot \left( \frac{W_{shmax}^{rf}-{W_{sh}}^{rf}}{rf\cdot W_{shmax}^{rf}} \right)$$ {#eq-dWshdt}

Where $r_{Ws}$ is a growth rate parameter, $T_{eff}$ is the effective temperature, $W_{shmax}$ is the maximum attained shoot dry matter and rf is form parameter. This function can be integrated numerically, using a value for $W_{sh}$ at sowing of 10 $g\cdot m^{-2}$. The effective temperature was calculated from:

$$T_{eff}=\max \left( 0,\left( {{T}_{a}}-{{T}_{b}} \right) \right) $$ {#eq-Teff}

Where $T_a$ is the daily average air temperature $T_b$ is a base temperature, assumed to be 4°C.

## Rooting depth

Rooting depth is in the submodel TSimpleRootModDM defined as the depth of the deepest root. It is mainly governed by temperature and the submodel is currently parameterized using air temperature as an input. Additionally soil texture and bulk density may optionally be used to modify the maximum rooting depth and the rate of rooting depth increase.

There are also two choices for modelling rooting depth increase: a simple linear increase with temperature sum and an expo-linear increase with temperature sum, which accounts for the observed lag phases in rooting depth increase during early development stages.

### Linear increase of rooting depth with temperature sum

Rooting depth zr (cm) is often found to increase linearly with accumulated temperature sum within certain development stages, but lag phases in rooting depth increase [@ThorupKristensen1998; @ThorupKristensen1998b] as well as diminishing rooting depth increases during maturity [@Jaafar1993] have been observed. Rooting depth increase $[cm \cdot d^{-1}]$ therefore simply is:

$$\frac{dz_r}{dt} =
\begin{cases}
0, & z_r > z_{r,\max} \\
k_{zb}\cdot T_{eff} \cdot f_{texture}, & \text{otherwise}
\end{cases}$$ {#eq-dzrdt}

Where $k_{zb}$ $[cm\cdot°C^{-1}\cdot d^{-1}]$ is a parameter. For many cereal crops it is about 0.1 $[cm\cdot°C^{-1}\cdot d^{-1}]$. The base temperature for root growth was set to 0°C.

The parameter $z_{r,\max}$ is the maximum rooting depth. Its value may be internally modified by the effects effects of soil texture and bulk density, which are considered by the correction factor $f_{texture}$ (see below).

### Expo-linear increase of rooting depth with temperature sum

The linear increase of rooting depth with temperature sum may be modified by an exponential term to account for the observed lag phases in rooting depth increase during early development stagese ([@thorup-kristensen1998][@ThorupKristensen1998c].

During the first phase rooting depth increase, dzr/dt (cm.d-1) therefore is:

$$ \frac{dz_r}{dt}=k_{za} \cdot T_{eff} \cdot z_r$$ {#eq-dzrdtExp}

Where T~eff~ (°C) is the effective temperature for root growth and k~za~ is a constant (°C^-1.^d^-1^).

This equation can be integrated:

$$z_{r}=z_{r0} \cdot {e^{{{k}_{za}}\cdot TS}}$$ {#eq-zrExp}

where z~r0~ is the rooting depth at the day of sowing/transplanting and TS the sum of average daily temperatures above the base temperature.

In order to obtain a continuously derivable function the exponential and the linear part of the function have to predict the same rooting depth increase at the switching point from one part to the other. Therefore, by combining the right hand sides of equation (7-7) and (7-9) and solving for zr at this switching point, zrc, where maximum rooting depth increase is reached can be calculated:

$${{z}_{rc}}=\frac{{{b}_{zr}}}{{{a}_{zr}}}$$ {#eq-zrc}

The temperature sum at which the rooting depth increase switches from the exponential to the linear phase, TScrit, is obtained by substituting zrc from Eqn. 7-10 for zr in Eqn. 7-8. Rearranging gives:

$$T{{S}_{crit}}=\frac{\ln \left( \frac{{{z}_{rc}}}{{{z}_{r0}}} \right)}{{{a}_{zr}}}$$ {#eq-TScrit}

### Effect of texture and bulk density on rooting depth

The german soil classification (KA5) gives values for a parameter called effective rooting depth (W~eff~). This is not equivalent to maximum rooting depth z~max~, W~eff~ is rather a value that gives a rough estimate of available soil water when multiplied by the plant available soil water content. The values of W~eff~ were used to implement a texture and bulk density dependent correction factor for rooting depth increase, $f_{texture}$, which is calculated as the ratio of effective rooting depth of the standard soil being an Ut3 texture with a bulk density class of 3. The values of W~eff~ for different soil textures and bulk density classes are shown in @tbl-TextureBDeffect.

```{r}
#| echo: false
#| message: false
#| warning: false
#| include: true
#| label: tbl-TextureBDeffect
#| tbl-cap: Effective rooting depth [dm] and relative effective rooting depth [-] for different soil textures and bulk density classes.


library(dplyr)
library(DT)
library(htmlwidgets)


fn <- "Components/Rootgrowth/Documentation/Weff.csv"
df.Weff <- read.csv(fn, header = TRUE, comment.char = "[", sep = ",", dec = ".")
names(df.Weff) <- c("Texture","BD", "Weff", "Weffrel")

num_cols <- names(df.Weff)[sapply(df.Weff, is.numeric)]



datatable(
  df.Weff,
  rownames = FALSE,
  filter = "top",
  options = list(
    pageLength = 8,
    autoWidth = TRUE,
    scrollX = TRUE
  ),
  callback = JS(
    "table.on('init.dt', function() {",
    "  var headerRows = $(table.table().header()).find('tr');",
    "  if (headerRows.length > 1) {",
    "    $(headerRows[1]).find('th').each(function(i) {",
    "      if (i > 1) {",
    "        $(this).html('');",
    "      }",
    "    });",
    "  }",
    "});"
  )
) %>%
  formatRound(columns = num_cols, digits = 3, dec.mark = ".")


```

```{r}
#| label: rooting-depth-static-interactive
#| echo: false
#| message: false
#| warning: false
#| include: true
#| fig-cap: "Effect of texture and bulk density on rooting depth dynamics"

library(dplyr)
library(tidyr)
library(plotly)
library(crosstalk)
library(htmltools)

# Assumptions:
# df.Weff contains columns Texture, BD, Weffrel

bzr <- 0.1
zrmax <- 120
tsum_max <- 2000

plot_df <- df.Weff %>%
  transmute(
    texture = Texture,
    bulkdensity = BD,
    Weffrel = Weffrel
  ) %>%
  filter(!is.na(texture), !is.na(bulkdensity), !is.na(Weffrel)) %>%
  distinct() %>%
  expand_grid(TSUM = seq(0, tsum_max, by = 10)) %>%
  mutate(
    zrmax_eff = zrmax * Weffrel,
    zr = pmin(zrmax_eff, TSUM * bzr * Weffrel),
    scenario = paste0(
      "Texture=", texture,
      ", BD=", bulkdensity,
      ", Weffrel=", round(Weffrel, 3)
    )
  )

sd <- SharedData$new(plot_df, group = "rooting")

bscols(
  widths = c(3, 9),
  list(
    filter_select("texture", "Texture class", sd, ~texture, multiple = TRUE),
    filter_select("bd", "Bulk density class", sd, ~bulkdensity, multiple = TRUE)
  ),
  plot_ly(
    sd,
    x = ~TSUM,
    y = ~zr,
    color = ~scenario,
    type = "scatter",
    mode = "lines",
    hovertemplate = paste(
      "TSUM=%{x}",
      "<br>zr=%{y}",
      "<extra>%{fullData.name}</extra>"
    )
  ) %>%
    layout(
      title = "Effect of texture and bulk density on rooting depth dynamics",
      xaxis = list(title = "Temperature sum (TSUM [°C d])"),
      yaxis = list(title = "Rooting depth zr (cm)", autorange = "reversed")
    )
)



```

## Vertical root distribution

Root length density, however, is not uniform over the soil profile. Typically it is decreasing exponentially from the highest values near the soil surface with depth [@Gerwitz1974; @Barraclough1984; @Greenwood1982]:

$$L_{rv}=L_{rv0}\cdot {{e}^{-{{k}_{r}}\cdot z}} $$ {#eq-Lrv}

where the constant $k_r$ $(cm^{-1})$ is the fractional decrease in RLD per unit increase of soil depth and $L_{rv0}$ is the root length density at zero soil depth.

Integration of z=0 to a depth $z=z_r$ were the root length density is very low yields the root length RL $(cm \cdot cm^{-2})$ [see @Kage2000]:

$$L_r=\int\limits_{z=0}^{z=z_r}{L_{rv0}\cdot {{e}^{-{{k}_{r}}\cdot z}}\cdot dz}=\frac{L_{rv0}}{{{k}_{r}}}\cdot \left( 1-{{e}^{-{{k}_{r}}\cdot {{z}_{r}}}} \right) $$ {#eq-Lr}

To calculate the average rooting density $\overline{L_{rv}}$ $(cm \cdot cm^{-3})$ within a certain soil layer located between two soil depths z1 and z2 the above equation may be set up for both depths.The difference between RL at z2 and z1 divided by the distance z2-z1 gives the desired value of $RLD_{av}$:

$$\overline{L_{rv}}=\frac{L_{rv0}\cdot \left( {{e}^{-{{k}_{r}}\cdot {{z}_{1}}}}-{{e}^{-{{k}_{r}}\cdot {{z}_{2}}}} \right)}{{{k}_{r}}\cdot \left( {{z}_{2}}-{{z}_{1}} \right)} $$ {#eq-avLrv}

At the moment $k_r$ and $L_{rv0}$ remain unknown parameters.

for the depth $z=z_r$ and introducing a new parameter, $r_{RLD}$, describing the ratio of RLD at $z_r$, $RLD_{zr}$, and RLD~0~ the following identity can be found for the parameter $k_r$:

$${{k}_{r}}=-\frac{\ln \left( \frac{RL{{D}_{zr}}}{RL{{D}_{0}}} \right)}{{{z}_{r}}}=-\frac{\ln \left( {{r}_{RLD}} \right)}{{{z}_{r}}} $$ {#eq-kr}

Thereby, the introduction of $r_{RLD}$ avoids the necessity of using an iterative solution of the above set of equations.

Knowing this value the root length density at the soil surface can be calculated:

$$RLD_{0}=RL\cdot k_{r}\frac{1}{1-e^{-{k_r}\cdot z_r}} $$ {#eq-RLD0}

For late sown winter wheat on a loess loam soil [@Kage2001] estimated values for $L_{rv0}$ of about 4 $cm \cdot cm^{-3}$ and $K_r$ of about 0.03.

```{r}
#| label: LrvExample
#| message: false
#| warning: false
#| include: false

# Root length density at depth 0
Lrv0 <- 4 # cm.cm-3

# depth Lrv decrease parameter 
kr <- 0.03

# sequence of depths
z <- seq(0, 150, 5)

# sequence of root length densities
Lrv <- Lrv0*exp(-kr*z)
#plot(Lrv, -z )
# integration over depth gives total root length per area
Lr <- Lrv0/kr*(1-exp(-kr*max(z)))
#cat("For these values of Lrv0 and kr the total root length is", format(Lr, digits=4), "cm/cm2.")

```

For these values of $L_{rv0}$ and $k_r$ the total root length is `r format(Lr, digits=4)` $cm\cdot cm^{-2}$.

```{r}
#| label: fig-LrvDepth
#| echo: false
#| fig-align: center
#| fig-cap: Root length density (Lrv) vs. depth
#| fig-height: 3.5
#| fig-width: 3


#plot(Lrv, -z , type = "l", ylab="depth [cm]", xlab = "Lrv [cm/cm3]")

illu <- data.frame(-z, Lrv)

minor_breaks <- rep(1:9, 21)*(10^rep(-10:10, each=9))
MyBreaks <- c(1, 10, 100, 1000, 10000)
m <- NULL
m <-ggplot(data=illu)
m <- m + geom_line(aes(x=Lrv, y=z),lwd=1, lty=1, color="brown")
m <- m + scale_colour_brewer(palette="Set1")
m <- m + scale_y_reverse()
m <- m + scale_fill_brewer(palette="Set1")
m <- m + scale_x_continuous(position = "top")
m <- m + xlab("Lrv [cm/cm³]") + ylab("depth [cm]")
#m <- set_theme(m,14)
m <- m + theme_bw(base_size = 14)
#ggsave(filename = "LRV_depth.png", plot=m, units="cm", dpi = 300, height=15, width=22, bg = "transparent")
m

```

## Root length density classes within soil layers

The vertical distribution described above assigns one mean root length density to each soil layer. `TSimpleRootModDM` can additionally represent spatial variation around this layer mean with ten equally weighted root length density (RLD) classes. These classes are named *moments* in parts of the source code, but their values are representative quantiles of a lognormal distribution rather than statistical moments. The feature is implemented for 20 soil layers and for both total RLD (`Wld_moments`) and effective RLD (`EffWld_moments`).

### Coefficient of variation

For layer $i$, the coefficient of variation $CV_i$ (%) is calculated from the mean total root length density $\overline{L}_{rv,i}$:

$$CV_i = \max\left[0,\; CV_{int} + CV_{slope}\ln\left(\overline{L}_{rv,i}\right)\right]$$ {#eq-rld-cv}

where $CV_{int}$ is `VK_lrv_int` and $CV_{slope}$ is `VK_lrv_slope`. The logarithm uses the numerical value of `Wld_arr[i]` in its declared unit of $cm\,cm^{-3}$. If the mean total RLD is zero or negative, $CV_i$ is set to zero. The same layer-specific $CV_i$ is used to construct the total and effective RLD classes.

### Construction of the ten classes

The percentage coefficient of variation is first converted to a fraction, $c_i=CV_i/100$. The parameters of the corresponding lognormal distribution are

$$\sigma_{\ln,i}=\sqrt{\ln\left(1+c_i^2\right)}$$ {#eq-rld-log-sd}

and

$$\mu_{\ln,i}=\ln\left(\overline{L}_{rv,i}\right)-\frac{1}{2}\sigma_{\ln,i}^{2}.$$ {#eq-rld-log-mean}

For class $j=1,\ldots,10$, the preliminary representative value is evaluated at the midpoint of an equally probable interval of the standard normal distribution:

$$L^{*}_{rv,i,j}=\exp\left[\mu_{\ln,i}+\sigma_{\ln,i}\Phi^{-1}\left(\frac{j-0.5}{10}\right)\right],$$ {#eq-rld-class-preliminary}

where $\Phi^{-1}$ is the inverse standard normal cumulative distribution function. Thus, the probabilities are 0.05, 0.15, ..., 0.95. Because ten discrete quantiles do not reproduce the analytical mean exactly, the implementation rescales them:

$$L_{rv,i,j}=L^{*}_{rv,i,j}\frac{10\overline{L}_{rv,i}}{\sum_{j=1}^{10}L^{*}_{rv,i,j}}.$$ {#eq-rld-class-rescaling}

Consequently, the arithmetic mean of the ten class representatives equals the layer mean:

$$\frac{1}{10}\sum_{j=1}^{10}L_{rv,i,j}=\overline{L}_{rv,i}.$$ {#eq-rld-class-mean}

The same procedure is applied to the effective mean root length density `EffWld_arr[i]`, using the $CV_i$ obtained from total RLD. If a supplied mean is zero or negative, all ten representatives of that distribution are set to zero.

The reported standard deviations `WLD_stdev_i` and `effWLD_stdev_i` are calculated from the discrete representatives:

$$s_{L,i}=\sqrt{\frac{1}{10}\sum_{j=1}^{10}\left(L_{rv,i,j}-\overline{L}_{rv,i}\right)^2}.$$ {#eq-rld-class-sd}

### Activation and coupling to soil water uptake

The RLD classes are updated at every model time step only when `CalcMoments` is `true`. The default parameter values `VK_lrv_int = 0` and `VK_lrv_slope = 0` produce ten identical class values, so non-zero variability parameters are needed to represent within-layer heterogeneity.

Assigning a `TSoilWaterModelR` component to the `RootedSoilWaterModel` property links the total and effective RLD class matrices directly to that soil water model. To use the classes for water uptake, configure the linked soil water model with `SinkTermMethod = MFP` and `WithClasses = true`; see the [TSoilWaterModelR documentation](../../Soil/Documentation/TSoilWaterModelR.qmd). The soil water model then calculates MFP-limited uptake separately for every RLD class, maintains a separate water amount for each class, and sums the ten class contributions to recover each layer's sink term. `MFPuptakedistribution` controls whether potential uptake is allocated empirically or in proportion to maximum flow.

| Item | Unit or values | Default | Role |
|---|---:|---:|---|
| `CalcMoments` | `false`, `true` | `false` | Recalculate total and effective RLD classes at every time step. |
| `VK_lrv_int` | % | 0 | Intercept $CV_{int}$ in @eq-rld-cv. |
| `VK_lrv_slope` | % | 0 | Slope $CV_{slope}$ in @eq-rld-cv. |
| `VK_lrv_i` | % | calculated | Coefficient of variation in layer $i$. |
| `WLD_stdev_i` | $cm\,cm^{-3}$ | calculated | Standard deviation of the ten total-RLD representatives in layer $i$. |
| `effWLD_stdev_i` | $cm\,cm^{-3}$ | calculated | Standard deviation of the ten effective-RLD representatives in layer $i$. |
| `RootedSoilWaterModel` | component reference | none | Links the RLD class matrices to a `TSoilWaterModelR` instance. |

## Boxcar representation of active root length

In `USimpleRootModDM.pas`, active root length is represented with a **boxcar ageing scheme**. These temporal age classes are distinct from the within-layer RLD classes described above. For each soil compartment, root length is stored in a matrix of discrete age classes (`Root_Matrix[i,j]`). Each column corresponds to one age class, so the matrix keeps track of how much root length in a compartment belongs to young, recently formed roots and how much belongs to older roots.

At the beginning of a simulation run, the age-class counter `n_age_cl` is set to zero. During the simulation it increases by one per day until the maximum number of age classes (`MaxAgeCl`) is reached. On each day, the content of the age classes is shifted by one position toward older classes:

$$RootMatrix_{i,j}(t) = RootMatrix_{i,j-1}(t-1)
\qquad \text{for } j = n\_age\_cl, \dots, 2$$ {#eq-boxcarShift}

This means that root length formed on previous days ages stepwise from one class to the next.

### Daily root increment

For each root compartment (i), total root length is obtained from root length density and compartment thickness:

$$WL_i(t) = WLD_i(t) \cdot \Delta z_i$$ {#eq-rootLength}

where

$$\Delta z_i = z_i - z_{i-1}$$ {#eq-compartmentThickness}

is the thickness of compartment (i). The newly formed root length for the current day is then computed as the positive increment:

$$\Delta WL_i(t) = \max \left(0,\; WL_i(t) - WL_i(t-1)\right)$$ {#eq-rootIncrement}

This increment is assigned to the youngest age class:

$$RootMatrix_{i,1}(t) = \Delta WL_i(t)$$ {#eq-newRootLength}

Thus, the first boxcar always contains the new roots formed during the current time step.

### Effective root length

Only the younger age classes are considered physiologically active. The effective root length in compartment (i) is therefore calculated as the sum over all age classes up to the active root duration:

$$EffWL_i(t) =
\sum_{j=1}^{\min(\mathrm{ActiveDuration},\,\mathrm{MaxAgeCl})}
RootMatrix_{i,j}(t)$$ {#eq-effectiveRootLength}

This means that roots older than `ActiveDuration` are still tracked in the boxcar matrix, but they no longer contribute to effective root activity.

The total effective root length in the whole rooted soil profile is then:

$$SRL_{eff}(t) = \sum_i EffWL_i(t)$$ {#eq-totalEffectiveRootLength}

### Effective root length density

For each compartment, the effective root length density is derived by dividing effective root length by compartment thickness:

$$EffWLD_i(t) = \frac{EffWL_i(t)}{\Delta z_i}$$ {#eq-effectiveRootLengthDensity}

This quantity represents the density of roots that are still considered active in water and nutrient uptake.

### Interpretation

The boxcar approach is a simple discrete ageing model. New roots enter the first age class, then move day by day into older classes. Root activity is restricted to the youngest fraction of the root system, as defined by `ActiveDuration`. In this way, the algorithm distinguishes between total root length and the smaller fraction of roots that are still effective for uptake processes.

## Parameterisation for silage maize

The calibration of the root growth parameters was done with independent data of the field trials with a newer hybrid only (KWS Stefano, field trials in 2021-22 at the experimental farm Hohenschulen of Kiel University (54.18°N, 9.58°E, 18 m a.s.l.). Root length (RL) and root length density (RLD) were monitored to a soil depth of 120 cm, either by soil pits or by soil cores. Soil cores were sampled at silking and grain filling and divided into 10 cm sections. After washing the samples, extracted roots were scanned using the Epson Perfection V850 Pro. The scans were analysed using RhizoVision Explorer v2.0.3 (Seethepalli et al. 2021). Rooting depth (zr) was investigated at four timepoints during the growing period of 2022 measuring the deepest visible root in the soil pit.

![Panel A: Root depth growth. Linear relation of root depth $z$ to thermal time ($\\mathrm{d}z_r/\\mathrm{d}t = k_{zr} \\cdot T_{\\text{eff}}$, $k_{zr} = 0.0703$). Panel B: Vertical root growth is described by an exponential function over soil depth $z$ ($\\mathrm{RLD} = \\mathrm{RLD}_0 \\cdot e^{k_{r} \\cdot z_r}$). Panel C: Specific root length (SRL) depicted as the regression between root dry matter \[g/m²\] ($\\mathrm{DM}_{\\text{root}}$) and root length (RL).](Figure_roots.png){#fig-Maize_roots}

```{r}
#| include: false


fn <- here("csv_Hume_Docu/Docu_Docu_MAIZE.csv")

df <- read.delim(fn, header = TRUE, sep = ";")|>filter(Submodel=="SimpleRootModDM1") |> filter( EntityType == "Parameter") |> dplyr::select(EntityName, Units, Value, Comment)|>distinct()|>
    mutate(Value=as.numeric(Value))
names(df) <- c("Parameter", "Units", "Value", "Description")
```

## Parameters
The parameters configured for `SimpleRootModDM1` in the silage maize model are shown in @tbl-Parameter.


```{r}
#| label: tbl-Parameter
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Parameter Values, Units and Description fitted for silage maize"

df$Value <- format(df$Value,
#  round(df$Value,4),
 #nsmall = 2,
# trim = T, 
 scientific = FALSE
)

kable(df, escape = FALSE,
      format = "html",
      caption = "Parameter",
      booktabs = TRUE
 )|>
   kable_styling(full_width = FALSE, position = "center",
                 htmltable_class = "table"
  ) |>
  kableExtra::row_spec(0, bold = TRUE)

```
